\begin{homeworkProblem}{Short Tree Proofs}

    Let $G=(V,E)$ be an undirected graph with $|V| \geq 1$.

    \begin{itemize}
        \item[A)] Prove that every connected component in an acyclic graph is a tree.
        \item[B)] Suppose $G$ has $k$ connected components. Prove that if $G$ is acyclic, then $|E|=|V|-k$.
        \item[C)] Prove that a graph with $|V|$ edges contains a cycle. 
    \end{itemize}

    \part

    A graph is a tree if and only if it is connected and acyclic. A connected component is a set of vertices such that all vertices in the set are connected, i.e., there is a path between all pairs of vertices in the set. If a graph is acyclic, i.e., it contains no cycles, then its connected components must also contain no cycles. Therefore, since a connected component of an acyclic graph is, by definition, connected and acyclic, it itself is a tree. 
    \\ \\
    \part

    If $G$ is acyclic, then its connected components are acyclic as well. From part A), these connected components are trees. A graph, or in this case, a connected component, is a tree if and only if it's connected and has $|V|-1$ edges. Since there are $k$ connected components in $G$, there are $|E| = |V|-k$ edges in the graph.

    \[
        |E|=\sum_{i=1}^k \left( |V_i|-1 \right) = \sum_{i=1}^k |V_i| - \sum_{i=1}^k 1 = |V|-k
    \]

    \part

    Assume that the graph with $|V|$ edges instead contains no cycles, i.e., it is acyclic. From part B), we then know that $|E|=|V|-k$, where $k$ is the number of connected components in the graph. Therefore, our assumption that the graph contains on cycles leads to a contradiction, as $|V| \neq |V|-k$ for all $k \geq 1$. Since there is at least one connected component in any graph (the graph itself), a graph with $|V|$ edges is not acyclic and contains a cycle. 
    
\end{homeworkProblem}