\begin{homeworkProblem}{Universal Preference}
    
    Suppose that preferences in a stable matching system are universal: all $n$ jobs share the preferences $C_1 > C_2 > \dots >  C_n$ and all candidates share the preferences $J_1 > J_2 > \dots > J_n$.

    \begin{itemize}
        \item[A)] What pairing do we get from running the algorithm with jobs proposing? Prove that this happens for all $n$.
        \item[B)] What pairing do we get from running the algorithm with candidates proposing? Justify your answer.
        \item[C)] What does this tell us about the number of stable pairings? Justify your answer.
    \end{itemize}

    \part
    
    The pairings we get from running the stable matching algorithm with jobs proposing is one where, for an index $i$, candidate $C_i$ is paired with job $J_i$, i.e., the pairings we get are $\{ (C_1, J_1), (C_2, J_2), \dots, (C_n, J_n) \}$.
    \\ \\
    To prove that this happens for all $n$, I'll use the principle of induction to prove the \textit{base case} and the \textit{inductive step}.

    \begin{itemize}
        \item \textit{Base Case}: When $n=1$, the claim holds since there is only one possible pairing $(C_1, J_1)$.
        \item \textit{Inductive Hypothesis}: Assume that the pairing we get is $\{(C_1, J_1), (C_2, J_2), \dots, (C_k, J_k)\}$ for some value of $n=k$ where $k \in \mathbb{N}$.
        \item \textit{Inductive Step}: For $n = k+1$, we can show that the preferences and matching are the following:

        \begin{table}[h!]
        \centering
        \renewcommand{\arraystretch}{1.3} % Adjust row height
        \setlength{\tabcolsep}{10pt} % Adjust cell padding

        \begin{tabular}{cc}
            \begin{tabular}{|c|c|}
            \hline
            \textbf{Jobs} & \textbf{Candidates} \\ \hline
            $J_1$ & $C_1 > C_2 > \dots > C_k > C_{k+1}$ \\ \hline
            $J_2$ & $C_1 > C_2 > \dots > C_k > C_{k+1}$ \\ \hline
            $\vdots$ & $\vdots$ \\ \hline
            $J_k$ & $C_1 > C_2 > \dots > C_k > C_{k+1}$ \\ \hline
            $J_{k+1}$ & $C_1 > C_2 > \dots > C_k > C_{k+1}$ \\ \hline
            \end{tabular}
            &
            \begin{tabular}{|c|c|}
            \hline
            \textbf{Candidates} & \textbf{Jobs} \\ \hline
            $C_1$ & $J_1 > J_2 > \dots > J_k > J_{k+1}$ \\ \hline
            $C_2$ & $J_1 > J_2 > \dots > J_k > J_{k+1}$ \\ \hline
            $\vdots$ & $\vdots$ \\ \hline
            $C_k$ & $J_1 > J_2 > \dots > J_k > J_{k+1}$ \\ \hline
            $C_{k+1}$ & $J_1 > J_2 > \dots > J_k > J_{k+1}$ \\ \hline
            \end{tabular}
        \end{tabular}
        \end{table}

        The algorithm takes $k+1$ days to produce a stable matching. The resulting pairing is as follows. The circles indicate the job that a candidate picked on a given day (and rejected the rest).

        \begin{table}[h!]
        \centering
        \renewcommand{\arraystretch}{1.3} % Adjust row height
        \setlength{\tabcolsep}{10pt} % Adjust cell padding
        \begin{tabular}{|c|c|c|c|c|c|}
            \hline
            \textbf{Candidate} & \textbf{Day 1} & \textbf{Day 2} & \dots & \textbf{Day $k$} & \textbf{Day $k+1$} \\ \hline
            $C_1$ & $\Circled{1}, 2, \dots, k, k+1$ & \Circled{1} & \dots & \Circled{1} & \Circled{1} \\ \hline
            $C_2$ &  & $\Circled{2}, \dots, k, k+1$ & \dots & \Circled{2} & \Circled{2} \\ \hline
            \vdots & \vdots & \vdots & $\ddots$ & \vdots & \vdots \\ \hline
            $C_k$ &  &  &  & $\Circled{k}, k+1$ & \Circled{$k$} \\ \hline
            $C_{k+1}$ &  &  &  &  & \Circled{k+1} \\ \hline
        \end{tabular}
        \end{table}

        It takes 1 more day for the algorithm to terminate, and since candidate $C_{k+1}$ is least preferred by all jobs, it will be proposed to last. The job proposing will be job $J_{k+1}$, since it is least preferred by all candidates. Therefore, we get a pairing $\{(C_1, J_1), (C_2, J_2), \dots, (C_k, J_k), (C_{k+1}, J_{k+1})\}$. By induction, we have shown that we get a pairing $\{ (C_1, J_1), (C_2, J_2), \dots, (C_n, J_n) \}$ for all $n$.
        
    \end{itemize}

    \pagebreak
    
    \part
    
    The pairings we get from running the stable matching algorithm with jobs proposing is one where, for an index $i$, candidate $C_i$ is paired with job $J_i$, i.e., the pairings we get are $\{ (C_1, J_1), (C_2, J_2), \dots, (C_n, J_n) \}$.
    \\ \\
    \part
    
    Because both the job-optimal and candidate-optimal stable matching algorithms produce the same pairing, there is only 1 stable pairing.
    
    
\end{homeworkProblem}