\begin{homeworkProblem}{Administrivia}
    \begin{itemize}
        \item[(a)] Make sure you are on the course Ed (for Q\&A) and 
        Gradescope (for submitting homeworks, including this one). Find and 
        familiarize yourself with the course website. What is its 
        homepage’s URL?
        \item[(b)] Read the policies page on the course website.
            \begin{itemize}
                \item[(i)] What is the breakdown of how your grade is 
                calculated, for both the homework and the no-homework 
                option? 
                \item[(ii)] What is the attendance policy for discussions?
                \item[(iii)] When are homeworks released and when are they 
                due? 
                \item[(iv)] How many ``drops" do you get for homeworks? How 
                many mini-vitamins will contribute to your grade? 
                \item[(v)] When is the final exam?
                \item[(vi)] What percentage score is needed to earn full 
                credit on a homework?
            \end{itemize}
    \end{itemize}

    \solution

    \begin{itemize}
        \item[(a)] \href{http://www.eecs70.org/}{http://www.eecs70.org/}
        \item[(b)] Policies:
            \begin{itemize}
                \item[(i)] Homework and no-homework grading:
                    \begin{multicols}{2}
                        \begin{itemize}
                            \item Discussion attendance: 5\%
                            \item Mini-Vitamins: 5\% (top 13 scores)
                            \item Homeworks: 15\% (lowest 3 dropped)
                            \item Midterm: 30\%
                            \item Final: 45\%
                        \end{itemize}
                        
                        \columnbreak % Forces a column break, if needed.
                        
                        
                        \begin{itemize}
                            \item Discussion attendance: 5\%
                            \item Mini-Vitamins: 5\% (top 13 scores)
                            \item Midterm: 36\%
                            \item Final: 54\%
                        \end{itemize}
                    \end{multicols}
                    
                    
                \item[(ii)] You need to attend at least 13 discussions to recieve full credit.
                \item[(iii)] Homeworks are released on Sunday and due Saturday of the next week at 4:00 pm.
                \item[(iv)] You get three drops for your homework. Only the top 13 scores for the Mini-Vitamins contribute towards your grade.
                \item[(v)] May 16, 2025 from 7:00 pm until 10:00 pm
                \item[(vi)] 73\%
            \end{itemize}
    \end{itemize}

\end{homeworkProblem}