\begin{homeworkProblem}{Course Policies}
    Go to the course website and read the course policies carefully. Leave 
    a followup on Ed if you have any questions. Are the following 
    situations violations of course policy? Write "Yes" or "No", and a 
    short explanation for each.

    \begin{itemize}
        \item[(a)] Alice and Bob work on a problem in a study group. They 
        write up a solution together and submit it, noting on their 
        submissions that they wrote up their homework answers together.
        \item[(b)] Carol goes to a homework party and listens to Dan 
        describe his approach to a problem on the board, taking notes in 
        the process. She writes up her homework submission from her notes, 
        crediting Dan.
        \item[(c)] Erin comes across a proof that is part of a homework 
        problem while studying course material. She reads it and then, 
        after she has understood it, writes her own solution using the same 
        approach. She submits the homework with a citation to the website. 
        \item[(d)] Frank is having trouble with his homework and asks Grace 
        for help. Grace lets Frank look at her written solution. Frank 
        copies it onto his notebook and uses the copy to write and submit 
        his homework, crediting Grace. 
        \item[(e)] Heidi has completed her homework using \LaTeX. Her 
        friend Irene has been working on a homework problem for hours, and 
        asks Heidi for help. Heidi sends Irene her PDF solution, and Irene 
        uses it to write her own solution with a citation to Heidi.
        \item[(f)] Joe found homework solutions before they were officially 
        released, and every time he got stuck, he looked at the solutions 
        for a hint. He then cited the solutions as part of his submission.
    \end{itemize}

    \textbf{Justification}
    \\

    \begin{itemize}
        \item[(a)] \textbf{Yes}, Alice and Bob must each write up their own 
        solution to the homework problems.
        \item[(b)] \textbf{No}, Carol wrote her own solution to the 
        homework problems and credited her source.
        \item[(c)] \textbf{No}, Erin wrote her own solution to the homework 
        problem and credited her source.
        \item[(d)] \textbf{Yes}, Frank must come up with his own solution 
        and not copy Grace's verbatim. 
        \item[(e)] \textbf{Yes}, Irene must come up with her own solution 
        and not copy Heidi's verbatim, even if it is a different medium.
        \item[(f)] \textbf{No?}
    \end{itemize}
    
\end{homeworkProblem}